---
title: "Income Prediction with Decision Trees"
output:
  html_document:
    df_print: paged
  html_notebook: default
  word_document: default
---

<h2> Business Understanding </h2>

Objective is to use information about existing customers of a financial services company to develop a model that predicts whether a customer has an income of $50,000 or more. The motivation for this problem is to identify potential high-income customers from a prospective customer database that was recently purchase

This problems is going to be solved using a classification tree

<h2> Data Understanding </h2>

```{r}
library(tidyverse)
income <- read_csv("Student/Data/income.csv", col_types = "nffnfffffnff")
glimpse(income)

```

<h3> Exploring the data </h3>

```{r}
summary(income)

```
<h2> Data Preparation </h2>

Splitting test and train data

```{r}
set.seed(1234)
sample_set <- sample(nrow(income), round(nrow(income) * .75), replace = FALSE)

income_train <- income[sample_set, ]
income_test <- income[-sample_set, ]
```

Review data set balance

```{r}
income %>%
  select(income) %>%
  table(exclude = NULL) %>%
  prop.table() %>%
  round(4) * 100
```

Review test data set balance

```{r}
income_test %>%
  select(income) %>%
  table(exclude = NULL) %>%
  prop.table() %>%
  round(4) * 100
```

Review train data set balance

```{r}
income_train %>%
  select(income) %>%
  table(exclude = NULL) %>%
  prop.table() %>%
  round(4) * 100
```

<h3> Balancing the Training  Data Set </h3>

```{r}

library(DMwR)
set.seed(1234)

income_train <- SMOTE(income ~ ., data.frame(income_train), perc.over = 100, perc.under = 200)

income_train %>%
  select(income) %>%
  table(exclude = NULL) %>%
  prop.table() %>%
  round(4) * 100

```

<h2> Modeling </h2>

Training the model

```{r}
library(rpart)

income_mod <- 
  rpart (
    income ~ ., 
    method = "class", 
    data = income_train
  )

```


<h2> Model Evaluation </h2>

Plot the Decision Tree

```{r}

library(rpart.plot)
rpart.plot(income_mod)

```

Confusion Matrix

```{r}

income_pred <-  predict(income_mod, income_test, type = "class")
income_pred_table <- table(income_test$income, income_pred)
income_pred_table
```

Prediction Accuracy

```{r}

sum(diag(income_pred_table)) / nrow(income_test)

```

























