---
title: "Case Study: Predict Book Pressure"
output: html_notebook
---


Load necessary libraries

```{r}
library(tidyverse)
```

Load health data set

```{r}
health <- read_csv("health.csv")
```

Preview Data

```{r}
glimpse(health)
```
Convert variables $diabetes, $smoker into factors

```{r}
health <- health %>%
  mutate(diabetes = as.factor(diabetes)) %>%
  mutate(smoker = as.factor(smoker))

```

<h3> EDA Section

``` {r}
summary(health)
```

Histogram: systoilic blood pressure - statistical distribution of the predicted variable

```{r}
health %>%
  ggplot() +
  geom_histogram(mapping = aes(x = systolic), fill = "lightblue", color = "black") +
  theme_minimal()
```

Statistical distribution of the predictor variables

```{r}
health %>%
  select(-systolic) %>%
  keep(is.numeric) %>%
  gather() %>%
  ggplot () +
    geom_histogram(mapping = aes(x=value,fill=key), color = "black") +
    facet_wrap(~ key, scales = "free") +
  theme_minimal()
```

Correlation analysis among peredictor variables

```{r}
cor(health[, c("systolic", "weight", "height", "bmi", "waist", "age", "fastfood")])

```

Visualize Correlations of Independent Variables

``` {r}
corrplot(health_correlations, type="upper")

```
Simple Linear regression model - single variable (age)

```{r}
health_mod1 <-  lm(data = health, systolic ~ age)

summary(health_mod1)
```
Multiple Linear Regression Model

``` {r}
health_mod2 <- lm(data = health, systolic ~ .)

summary(health_mod2)

```
Run Test for zero mean of residuals

```{r}
mean(health_mod2$residuals)

library(olsrr)

ols_plot_resid_hist(health_mod2)

```
Check Presence of Heterodasticity

```{r}
ols_plot_resid_fit(health_mod2)
```
Run test for residual autocorrelation {Durbin-Watson Statistic}

```{r}
library(car)

durbinWatsonTest(health_mod2)


```

Check for influencial points in health data (outliers)

```{r}
ols_plot_cooksd_chart(health_mod2)

```

Review statistical distribution of the outliers

```{r}
outlier_index <- as.numeric(unlist(ols_plot_cooksd_chart(health_mod2, print_plot = FALSE)$outliers[,"observation"]))
outlier_index
```

Compare Summary Statistics of Outliers

``` {r}
summary(health[outlier_index,])

```

Outliers Removed

```{r}
summary(health[-outlier_index,])
```

Remove Outliers from Dataset

```{r}
health2 <- health[-outlier_index,]
```

Test for multicollinearity - (VIF above 5.0 indicates multicollinearity)

```{r}
ols_vif_tol(health_mod2)
```

Build a model accounting for multicollinearity

```{r}
health_mod3 <- lm(data=health2, systolic ~ weight+age+diabetes)

summary(health_mod3)
```
Introduce non-linear predictors in the model

```{r}
health2 <- health2 %>%
  mutate(age2 = age^2,
         lage=log(age))

```

Build model with ols_step_both_p() function from olsrr package

```{r}

ols_step_both_p(
  model = lm (
    data = health2,
    systolic ~ weight * diabetes + age * diabetes + age2 * diabetes + lage * diabetes
  ),
  pent = 0.2,
  prem = 0.01,
  details = TRUE
)

```






















